---
title: "Specifying Recruitment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Specifying Recruitment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

As part of the operating model (OM) list object (`om`) that is a required input to the MSE simulation function, users must specify a recruitment function as well as relevant parameter values for the OM to project future recruitment. Some simple recruitment functions are prepackaged with the model -- a Beverton-Holt stock recruit function, a function that resamples from a historical recruitment timeseries, and a function that specifies distinct recruitment regimes -- but users are also able to specify their own custom recruitment functions, as necessary. 

### The Recruitment Object
The recruitment list object requires two components: (1) a reference to an R function that specifies how future recruitment is to be generated (the "recruitment function"), and (2) a list of parameter values to pass to said function (the "recruitment parameters").

The recruitment function must return either a single value, indicating the total recruitment for the population across all spatial regions, or a vector/array of values indicating the recruitment to each spatial region (should be dimension `[1, nregions]`). If a single, system-wide, recruitment value is returned, additional steps are required to define how to apportion recruits to the different spatial regions in the model (a "recruit apportionment method").

In the most simple case (below), future recruitment is generated from the `recruit_func` function, which simply returns a recruitment of "10" for each of the 5 regions for the next 100 years.

```{r, recruitment_obj, eval=FALSE}
om <- readRDS(file.path("data", "spatial_sablefish_om.RDS"))

recruit_func <- function(nyears, nregions, seed){
    set.seed(seed)
    return(array(10, dim = c(nyears, nregions)))
}

recruitment_obj <- list(
    # function from which to generate future recruitment events
    func = recruit_func,
    # extra parameters to pass to `func`
    pars = list(
        nyears = 100,
        nregions = 5
    )
)

om$recruitment <- recruitment_obj
```
Note that the `seed` parameter is a required input to the recruitment function, but does not need to be specified in the `recruitment$pars` list object. The MSE simulation loop will automatically pass the appropriate seed value to the recruitment function internally.

For recruitment functions that return only a single value, two additional entries in the recruitment list object must be provided to specify how to apportion recruits to the different spatial regions in the model: (1) a reference to an R function that specifies how recruits should be apportioned (the "recruit apportionment function"), and (2) a list of parameters values to pass to said function (the "recruit apportionment parameters"). The example below defines a recruit apportionment functions that divides the total recruitment evenly across all spatial regions every year:

```{r, recruit_apportionment, eval=FALSE}
nregions <- 5

recruit_apportionment_func <- function(nregions, seed){
    return(array(1 / nregions, dim = c(1, nregions)))
}

om$recruitment$apportionment <- list(
    # function from which to apportion recruits
    func = recruit_apportionment_func,
    # extra parameters to pass to `func`
    pars = list(
        nregions = nregions
    )
)

```
Note that the `seed` parameter is a required input to the recruit apportionment function, but does not need to be specified in the `recruitment$apportionment$pars` list object. The MSE simulation loop will automatically pass the appropriate seed value to the recruit apportionment function internally. Also note that recruit apportionment functions should only return a vector/array of dimension `[1, nregions]`. Recruit apportionment should not be calculated for all years up front, but rather for each year as the MSE simulation loop runs.

Recruitment functions can be as complicated as necessary, and can also make use of annual population state variables, such as spawning biomass. Examples of more complex recruitment functions are available at the end of this page.

### Provided Recruitment Functions

Four generic recruitment functions are provided as part of the `SpatialSablefishMSE` package. They cover four common ways to simulate future fisheries recruitment.

##### Resample Recruitment
The first provided recruitment function generates future recruitment by resampling from an array of provided values. Most often, the array being resampled from is an array of historical spatial recruitment events, but can also include other values. 

```{r, resample_recruitment, eval=FALSE}
resample_recruits_spatial <- function(hist_recruits, nyears, seed){
    set.seed(seed)
    idx <- sample(1:nrow(hist_recruits), size=nyears, replace=TRUE)
    return(hist_recruits[idx,])
}

om$recruitment$func <- resample_recruits_spatial
om$recruitment$pars <- list(
    # array of values to resample from
    hist_recruits = hist_recruits,
    # number of years to generate future recruitment for
    nyears = nyears
)
```
Because this recruitment function returns a recruitment value for each spatial region, no additional recruit apportionment function is required.

##### Beverton-Holt Recruitment
The second provided recruitment function generates future recruitment via a Beverton-Holt stock recruitment relationship (Beverton and Holt, 1957), parameterized using steepness. The Beverton-Holt function is defined as:

$R_y = \frac{4hR_0\text{SSB}_{y-1}}{(1-h)S_0 + (5h-1)\text{SSB}_{y-1}}$

where $h$ is steepness, $S_0$ is unfished spawning biomass, $R_0$ is unfished recruitment, and $\text{SSB}$ is the current spawning stock biomass. Lognormally distributed deviations from the base stock recruit relationship, with recruitment variation $\sigma_R = 1.20$, are also applied. 

```{r, bevholt_recruitment, eval=FALSE}
om$recruitment$func <- beverton_holt
om$recruitment$pars <- list(
    h = 0.85,   # steepness
    R0 = 15,    # unfished recruitment
    S0 = 300,   # unfished spawning biomass
    sigR = 1.20 # recruitment variability
)

resample_recruit_apportionment <- function(hist_recruits, seed){
    set.seed(seed)
    idx = sample(1:nrow(hist_recruits), size=1, replace=TRUE)
    props <- t(apply(hist_recruits, 1, function(x) x/sum(x)))
    return(props[idx,])
}

om$recruitment$apportionment$func <- resample_recruit_apportionment
om$recruitment$apportionment$pars <- list(
    hist_recruits = hist_recruits
)
```
This recruitment function returns a single value for total recruitment, which is then apportioned to each spatial region using the `resample_recruit_apportionment` function. 

Annual spawning biomass (SSB) is passed to the underlying `beverton_holt` function automatically inside the MSE simulation loop. 

##### Regime Recruitment
The third provided recruitment functions generate future recruitment in distinct alternating regimes parameterize by different Beverton-Holt functions. Users have the ability to control the length, order, and relative strength of the two regimes. The recruitment function works by parameterizing two separate Beverton-Holt stock recruit relationships, differing in their value of unfished recruitment ($R_0$).  

```{r, regime_recruitment2, eval=FALSE}
om$recruitment$func <- bevholt_regimes
om$recruitment$pars <- list(
    # steepness
    h = 0.85,
    # spawning biomass per recruit (to calculate regime specific S0)
    sbpr = 20,                
    # Regime specfic R0
    R0 = c(12.5, 50),           
    # Regime specific recruitment variability
    sigR = c(1.20, 1.20),       
    # Total number of years to generate future recruitment for
    nyears = nyears,            
    # Lengths of each regime
    regime_length = c(20, 5),   
    # Start with first regime
    starting_regime = 0
)
om$recruitment$apportionment$func <- resample_recruit_apportionment
om$recruitment$apportionment$pars <- list(
    hist_recruits = hist_recruits
)
```
Like with the base Beverton-Holt recruitment function, this recruitment function returns a single value for total recruitment, which is then apportioned to each spatial region using the `resample_recruit_apportionment` function.

##### Crash Recruitment
The final provided recruitment function generates future recruitment through resampling, as in the "Resample Recruitment" option, but with a defined period where recruitment is different. This different period can be used to simulate a period depressed recruitment (e.g. a recruitment crash) or a period of amplified recruitment (e.g. a recruitment boom), depending on how it is parameterized. Users can specify the simulation year in which the alternative recruitment period begins, the length of that period, and the mean value of recruitment during that period. During the alternative recruitment period, lognormal deviations about the specified mean recruitment are applied using a $\text{CV}=0.1$.

```{r, crash_recruitment, eval=FALSE}
om$recruitment$func <- recruits_crash
om$recruitment$pars <- list(
    # Start a crash period in simulation years 1
    crash_start_year = 1,
    # 20-year crash period
    crash_length = 20,
    # Use minimum historical recruitment as average crash recruitment level
    crash_value = min(hist_recruits),
    # recruitment vector to resample from outside of the "crash period"
    hist_recruits = hist_recruits,
    # Total number of years to generte future recruitment for
    nyears = nyears
)
om$recruitment$apportionment$func <- resample_recruit_apportionment
om$recruitment$apportionment$pars <- list(
    hist_recruits = hist_recruits
)
```

### Custom Recruitment Functions

##### Required Parameters
The only required parameter for custom recruitment functions is a `seed` parameter, corresponding to a random seed. This is used to ensure that future recruitment vectors are reproducible across simulations. Note that while this `seed` parameter is a required input to the function, it does not need to be specified in the `recruitment$pars` list object. Instead, the MSE simulation loop will automatically pass the appropriate seed value to the recruitment function internally. 

##### State-Independent Recruitment Functions
State-independent recruitment functions are those that do not require knowledge of the current state of the population to generate future recruitments. These functions have the benefit of being able to generate all future recruitment events before entering the simulation loop. The `resample_recruitment` and `crash_recruitment` functions are examples of such state-independent recruitment functions.

As a extension of the provided `resample_recruits` functions, we provide the following example. Here, future recruitment is randomly resampled from a vector of historical recruitment events, with some weights applied, creating a new vector of recruitment for the next 100 years.

```{r, recruit_func2, eval=FALSE}
weighted_resample_recruitment <- function(hist_recruits, nyears, weights, seed){
    set.seed(seed)
    r <- sample(hist_recruits, nyears, replace=TRUE, prob=weights)
    return(r)
}


recruitment_obj <- list(
    # function from which to generate future recruitment events
    func = recruit_func,
    # extra parameters to pass to `func`
    pars = list(
        hist_recruits = seq(1, 100, 5)
        nyears = 100,
        weights = c(rep(0.025, 10), rep(0.05, 10))
    )
)

om$recruitment <- recruitment_obj

```

##### State-Dependent Recruitment Functions
State-dependent recruitment functions are those that do require knowledge of the current state of the population to appropriately generate future recruitment. These types of recruitment functions commonly require knowledge of annual SSB to parameterize some stock-recruit relationship, such as the Beverton-Holt or Ricker curves. The provided `beverton_holt` and `bevholt_regimes` functions are good examples of such state-dependent recruitment functions.   

The primary difference between defining state-dependent and state-independent recruitment functions is that state-dependent functions must be written as a ["function factory"](https://adv-r.hadley.nz/function-factories.html). Function factories are functions that parameterize, and then return a reference too, another function. Here, a function factory is used so that the parameter values of a stock recruit relationship can be defined by the user external to the simulation loop, and then the model can calculate annual SSB internally and pass that along to the output of the function factory. Additionally, many of these types of recruitment functions will require defining a recruit apportionment function as part of the recruitment object.

Below is an example of creating a new Beverton-Holt style stock recruit relationship, this time parameterized using $\alpha$ and $\beta$:

```{r, bevhlt2, eval=FALSE}
beverton_holt_ab <- function(a, b, seed){
    set.seed(seed)
    function(ssb, y){
        bh <- (a*ssb)/(1+b*ssb)
        return(bh)
    }
}

recruitment_obj <- list(
    # reference to the beverton_holt_ab function above
    func = beverton_holt_ab,
    # parameter values for 'a', and 'b'
    pars = list(
        a = 10
        b = 5
    )
)

om$recruitment <- recruitment_obj

# Add recruit apportionment function as well

```

The MSE simulation loop will handle calculating annual SSB and passing that value to the function factory internally. 

Note the lack of any variability around the stock-recruit relationship in the above example. For state-dependent recruitment functions, recruitment stochasticity is calculated and applied separately by the MSE simulation loop. Annual recruitment deviations are distributed as $~\mathcal{N}(0, \sigma_R)$ and are applied multiplicatively to the output from the above function.