---
title: "Processing MSE Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Processing MSE Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `run_mse()` and its associated wrapper functions return large quantities of data, inlcuding:

* Landed catch-at-age
* Fishing mortality-at-age
* Numbers-at-age
* Recommended F from the harvest control rule
* Computed TACs
* etc.

These data objects are all multi-dimensional arrays of dimensions `[nyears, nages, nsexes, nregions, nfleets, nsims]`.

Owing to the large number of outputs, their complex dimensions, and the frequency with which multiple MSE simulations  are compared, processing MSE results can be challenging. The `SablefishMSE` package provides a specific helper function for coercing output MSE data into a form that is useable for deriving additional quantities (e.g. spawning stock biomass), computing performance metrics, and plotting results.

#### `bind_mse_results`

The function `bind_mse_results(model_runs, var, extra_columns)` takes a list of completed MSE simulation objects, a vector of variable names (as characters), and a named list of additional columns, and returns a dataframe in long-format. It is intended to be used as below:

```{r, introduce_bind, eval=FALSE}
mse1 <- run_mse(om, hcr, mse_options, nyears_input=100)

model_runs <- list(mse1)
extra_columns <- list(om="hcr1")
vars <- c("naa")

results <- bind_mse_outputs(model_runs, vars, extra_columns)

```

The resulting object will be a data frame with at least six columns:

* time
* age
* sex
* region
* value
* L1

_Note that the "L1" column will contain the name(s) of the variable(s) each piece of data ("value") is associated with. In the above example, the "L1" column will be "naa"._

If `extra_columns` are specified, additional columns will be appended to the end of the dataframe. If mse object(s) in `model_runs` contain results from multiple simulations, a "sim" column will also be a part of the dataframe. Any combination of output variables can be supplied as part of the "vars" parameter vector, though the resulting dataframe may contain `NA` entries if mixing between variables that are defined "at-age" (e.g. numbers-at-age or F-at-age), and variables that are defined annualy (e.g. TAC or recommended F from the HCR function).

From here, users may interact with the MSE outputs with standard R dataframe commands, or via typical tidyverse functions. 

#### How to Use the `extra_columns` Parameter
The `extra_columns` parameter doesn't make a lot of sense in the above the example, as only a single MSE object is being processed. However, in the case where multiple MSE objects, with different combinations of OMs and HCRs are being processed simultaneously, users need a way to differentiate between which data belongs to which MSE simualtions. The `extra_columns` parameter provides a means for users to explicitly identfiy between MSE simulations.

```{r, extra_columns, eval=FALSE}
mse1 <- run_mse(om1, hcr1, mse_options, nyears_input=100)
mse2 <- run_mse(om1, hcr2, mse_options, nyears_input=100)
mse3 <- run_mse(om2, hcr1, mse_options, nyears_input=100)
mse4 <- run_mse(om2, hcr2, mse_options, nyears_input=100)

model_runs <- list(mse1, mse2, mse3, mse4)
extra_columns <- expand.grid(
    om = c("om1", "om2"),
    hcr = c("hcr1", "hcr2")
)
vars <- c("naa")

results <- bind_mse_outputs(model_runs, vars, extra_columns)
```
In the above example, four MSE simulations, spanning two OMs and two HCR functions, are all being processed together (presumably for comparing the performance of the different HCRs across both OMs later on). The `extra_columns` data.frame defines that two additional columns, "om" and "hcr", should be added to the resulting output dataframe.

#### Additional Helper Functions
Three helper functions are provided for quickly computing common derived quantities:

* `get_ssb_biomass()` - computes spawning biomass and total biomass from the true and estimated numbers-at-age matrices
* `get_fishing_mortalities()` - computes the fully selected fishing mortality for each fleet and jointly across all fleets from the true and estimated F-at-age matrices
* `get_recruits()` - computes annual recruitment from the true and estimated numbers-at-age matrices
* `get_landed_catch()` - computes total and fleet-specific annual landed catch
* `get_management_quantities()` - computes annual ABC, TAC, and expected landings (TAC * attainment). Historical management data from Goethel et al. 2023.
* `get_reference_points()` - computes F40, B40, and B0 reference points for all combinations of OMs and HCRs

The resulting tibble from each of these function remains in long format with the appropriately columns retained. For plotting, additional processing is likely needed.

---

_NOTE: Additional helper functions will likely be added throughout the lifecycle of this package to streamline common calculations. A `get_landings()` function is in development to allow simplified processing of landings data._
