---
title: "Running the MSE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the MSE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Once an operating model (OM) object, a recruitment object, and a management procedure (MP) object have all been appropriately defined, the MSE simulation loop can be almost be initiated. 

### MSE Options
A list of MSE options must be defined to control internal aspects of the MSE simulation loop, primarily how long to run the MSE for. The MSE options object requires the following inputs:

- `n_proj_years` - the number of projection years to run the MSE for. This is the number of years for which the specified management procedure will be applied.
- `n_spinup_years` - the number of years for which to apply the specfied OM before, applying the specified MP. For sablefish, this parameter is the number of years since 1960 (0-64), and indicates the calendar year you would like to begin MSE siulations from.
- `recruitment_start_year` - which year of the simulation new recruitment events should begin in. Often this is the same as `n_spinup_years` but can be longer if the user wants to use a specific recruitment timeseries at the beginning of the MSE simulations.
- `run_estimation` - whether the estimation method should run or not. If `FALSE` the management procedure will be applied to OM outputs directly rather than to estimates from the EM. 

Sensible defaults are available via the `setup_mse_options` function. Using them will begin the MSE simulations from the estimated population state in 2024, and project forward for 100 years.

```{r, source, include=FALSE}
source(file.path(here::here(), "R/setup_mse_options.R"))
```

```{r, mse_options2, eval=TRUE, echo=TRUE}
options <- setup_mse_options()
options
```

Note that the total number of simulation years for a single MSE run is `n_spinup_years + n_proj_years`.

### Parallel MSE Simulations
Frequently, it is desirable to run multiple MSE simulations across different random seeds, to evaluate an MP in the face of uncertainty. 

The `run_mse_parallel(...)` function is designed to act as a simple parallel wrapper for the base `run_mse(...)` function that will run multiple MSE simulations, with the same OM and MP specifications, where each simulation will only vary by the random seed used to generate annual recruitment levels and simulate observations.

To use `run_mse_parallel` users must specify the number of simulations to run, and provide a list of simulation seeds in addition to the OM, MP, and options objects that are required for the base `run_mse` function. `run_mse_parallel` will handle parallelizing the MSE simulations and aggregating the output data as simulations complete.

```{r, run_mse_parallel, eval=FALSE, echo=TRUE}
run_mse_parallel(
    nsims = 10,
    seeds = sample(1:1000, 10),
    om=om,
    hcr=mp,
    mse_options=options,
    nyears = 1
)
```

Note that this function will use up to `N-2` available compute cores on a machine, so as to allow the machine to remain functional. Parallel processing is enabled through use of the `pblapply` package and associated function. Users __DO NOT__ need to handle creating their own parallel processing cluster; this is handled internally.

### Multiple MSE Simulations
Often, multiple combinations of MPs and OMs are tested together within an MSE to evaluate MPs across different states of nature. A wrapper function `run_mse_multiple` is provided to faciliate this and return appropiately formatted MSE outputs.

To use `run_mse_multiple`, users should define multiple OM, MP, and/or MSE options objects, place them into lists, and pass the respective lists to `run_mse_multiple`:

```{r, run_mse_multiple, eval=FALSE}
run_mse_multiple(
    # List of OMs to run
    om_list = list(om1, om2),
    # List of MPs to run
    hcr_list = list(mp1, mp2),
    # List of MSE options to run
    mse_options_list = list(options)
    seed_list = sample(1:1000, 10),
    nyears=1
)
```

`run_mse_multiple` will run all combinations of OMs, MPs, and MSE Options provided in their respective lists, each for the number of simulations provided in `seed_list`. 